### 2017-03-10 real/draft.tex
[draft] tune last subsection

Let's consider the extreme case $\npack \to \infty$ so we can neglect
the uncertainty on $\alpha, \beta$ and
$p(\alpha, \beta \cond \npack, \bml) =
\delta(\alpha-\alpha_0)\delta(\beta - \beta_0)$ such that
\begin{align}
  \lleq{gmj}
  p(Q\cond n, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond n) p(Q\cond N, \alpha_0, \beta_0) \, .
\end{align}
For illustration only, let us consider the Poisson rate $\lambda$ were
known. Then learning $n$ would not improve our knowledge of $\lambda$
and we would have
\begin{align}
  \lleq{gmk}
  p(Q\cond \lambda, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond \lambda) p(Q\cond N, \alpha_0, \beta_0) \, ,
\end{align}
where $p(N \cond \lambda)$ is the Poisson distribution \refeq{pmb}.
In \reffig{comp-unc}, we illustrate the impact of the uncertainty due
to $\lambda$ or $\alpha, \beta$ for $n=0,2, 20$ packets and continue
to use as true values $\alpha_0 = 1.5, \beta_0 = 60$. For $n>0$, we
set $\lambda=n$ as the mode of the posterior $p(\lambda \cond n)$, and
for $n=0$, we set $\lambda = 0.347$ by solving
\begin{align}
  \lleq{gml}
  p(N=0 \cond \lambda) &= p(N=0 \cond n=0)
\end{align}
such that the contribution at $N=0$ is identical for the Poisson
likelihood and Poisson posterior predictive distribution.

It is interesting to note that the Jeffreys prior for $\alpha, \beta$
leads to an extra mode at $\Lumtot = 0$ visible in the case $n=2$. We
checked that this is true for every individual term in the
\refeq{gmhb}, and the superposition of $p(\Lumtot \cond N, n, \bml)$
for various $N$ further enhances this effect. The reason for the
boundary mode is that $\GammaDist(\Lumtot \cond \alpha, \beta)$ peaks
at the boundary for $\alpha<1$, and peaks away from $\Lumtot=0$ for
$\alpha>1$. As $n$ increases, the likelihood concentrates around
$\alpha=1.5$, hence the distribution becomes unimodal.

Already for $n=20$, it becomes apparent in the right panel of
\reffig{comp-unc} that the uncertainty for $\Lumtot$ is dominated by
the Poisson uncertainty: both $p(Q \cond n, \alpha_0, \beta_0)$ and
$p(Q \cond n, \bml)$ yield a standard error of $\approx 0.19$. This is
necessarily larger that the standard error of
$p(\Lumtot \cond \lambda, \alpha_0, \beta_0)$ because the variance of

The case where all parameters are known is not attainable in practice
and only shown for reference. But the case where $\alpha, \beta$ are
fixed can be approximated easily if many packets from other bins can
be combined.

Note that $n=2$ is the minimum number of packets to perform
the binned analysis.
\fred{I was playing around in mathematica and it seems even for $n=1$ the posterior has a well defined mode. I don't if it is proper, though. In any case it has huge variance.}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{comp_unc}
  \caption{Prediction of $\Lumtot$ for $n=0,2,20$ considering $\lambda, \alpha, \beta$ fixed (green), $\alpha, \beta$ fixed (red), or nothing fixed (blue).}
  \label{fig:comp-unc}
\end{figure}

@@

\subsection{Moments of the luminosity prediction}\label{sec:predict-moments}
\fred{Shouldn't we use $n, \bml$ instead of $\bmy$ to be consistent?
  $\bmy$ would mean we use all samples from all bins}

>>>>

Let's consider the extreme case $\npack \to \infty$ so we can neglect
the uncertainty on $\alpha, \beta$ and
$p(\alpha, \beta \cond \npack, \bml) =
\delta(\alpha-\alpha_0)\delta(\beta - \beta_0)$ such that
\begin{align}
  \lleq{gmj}
  p(\Lumtot \cond n, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond n) p(Q\cond N, \alpha_0, \beta_0) \, .
\end{align}
For illustration only, let us consider the Poisson rate $\lambda$ were
known. Then learning $n$ would not improve our knowledge of $\lambda$
and we would have
\begin{align}
  \lleq{gmk}
  p(\Lumtot \cond \lambda, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond \lambda) p(Q\cond N, \alpha_0, \beta_0) \, ,
\end{align}
where $p(N \cond \lambda)$ is the Poisson distribution \refeq{pmb}.
In \reffig{comp-unc}, we illustrate the impact of the uncertainty due
to $\lambda$ or $\alpha, \beta$ for $n=0,2, 20$ packets and continue
to use as true values $\alpha_0 = 1.5, \beta_0 = 60$. For $n>0$, we
set $\lambda=n$ as the mode of the posterior $p(\lambda \cond n)$, and
for $n=0$, we set $\lambda = 0.347$ by solving
\begin{align}
  \lleq{gml}
  p(N=0 \cond \lambda) &= p(N=0 \cond n=0)
\end{align}
such that the contribution at $N=0$ is identical for the Poisson
likelihood and Poisson posterior predictive distribution.

It is interesting to note that the Jeffreys prior for $\alpha, \beta$
leads to an extra mode at $\Lumtot = 0$ visible in the case $n=2$. We
checked that this is true for every individual term in the
\refeq{gmhb}, and the superposition of $p(\Lumtot \cond N, n, \bml)$
for various $N$ further enhances this effect.  The peak of
$\GammaDist(\Lumtot \cond \alpha, \beta)$ is at $\Lumtot=0$ for
$\alpha<1$, and at $\Lumtot > 0$ for $\alpha>1$. The Jeffreys prior
for $\alpha$ gives mass to both regions but as $n$ increases, the
likelihood concentrates around $\alpha=1.5$, hence the distribution
becomes unimodal.

For $n=20$, the right panel of \reffig{comp-unc} shows that
$p(Q \cond n, \bml)$ and $p(Q \cond n, \alpha_0, \beta_0)$ are very
similar so our reconstruction works correctly. While there is a small
shift in location, the variances are very similar:
$\variance{Q \cond n, \alpha_0, \beta_0}=0.036 \approx \variance{Q
  \cond n, \bml} = 0.034$.  This implies that the uncertainty due to
$\alpha, \beta$ is almost negligible compared to the effect of the
Poisson process. The simple explanation is that there is only one
observation to infer $\lambda$ but $n$ to infer $\alpha, \beta$ so for
large $n$ the Poisson uncertainty is dominant.

If $\lambda$ is known, the variance is signficantly smaller because
for the Poisson, $\variance{N \cond \lambda} = \lambda$, and for the
Negative Binomial $\variance{N \cond n} = 2(n + a-1)$, so for large
$n$, we have
\begin{align}
  \lleq{gmm}
  \variance{N \cond n} \approx 2 \variance{N \cond \lambda} \, ,
\end{align}
and this roughly explains
$\variance{Q \cond \lambda, \alpha_0, \beta_0 } = 0.021$.  The case
where all parameters are known is not attainable in practice and only
shown for comparison. But the case where $\alpha, \beta$ are fixed can
be reached easily if many packets from other bins can be
combined. Note that $n=2$ is the minimum number of packets to perform
the binned analysis.
\fred{I was playing around in mathematica and it seems even for $n=1$ the posterior has a well defined mode. I don't if it is proper, though. In any case it has huge variance.}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{comp_unc}
  \caption{Prediction of $\Lumtot$ for $n=0,2,20$ considering
    $\lambda, \alpha, \beta$ fixed (green), $\alpha, \beta$ fixed
    (red), or nothing fixed (blue). All densities are properly
    normalized to one.}
  \label{fig:comp-unc}
\end{figure}

@@

\subsection{Moments of the luminosity prediction}\label{sec:predict-moments}
\fred{Shouldn't we use $n, \bml$ instead of $\bmy$ to be consistent?
  $\bmy$ would mean we use all samples from all bins}
\fred{We need a reference to the appendix in the main text.}
